---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(dplyr)
v_zipfile <- "activity.zip"
v_datafile <- "activity.csv"
if (!file.exists(v_datafile)) {
        unzip(v_zipfile)
}
activity <- read.csv(v_datafile)
activity <- mutate(activity, date = as.Date(date,format="%Y-%m-%d"))

```


## What is mean total number of steps taken per day?
```{r}
stepsday <- summarize(group_by(activity,date), 
                      step_sum = sum(steps,na.rm=TRUE))
hist(stepsday$step_sum,xlab="Total steps per day",
     main = "Histogram of total number of steps taken each day")

steps_mean <- mean(stepsday$step_sum)
steps_mean_p <- round(steps_mean,1)
steps_mean_p

steps_median <- median(stepsday$step_sum)
steps_median_p <- round(steps_median,1)
steps_median_p
```
The mean of the total number of steps taken per day is: `r steps_mean_p`

The median of the total numbre of steps taken per day is `r steps_median_p`

## What is the average daily activity pattern?
```{r}
stepsinterval <- summarize(group_by(activity,interval), 
                      step_avg = mean(steps,na.rm=TRUE))
plot(stepsinterval$interval,stepsinterval$step_avg,type="l",
     xlab="interval (5 minute)", ylab = "Steps",
     main="Average number of steps taken")

max_steps <- max(stepsinterval$step_avg)
stepsinterval_fil <- filter(stepsinterval,step_avg==max_steps)
max_steps_interval <- stepsinterval_fil$interval
max_steps_interval
```
Which 5-minute interval, on average across all the days in the dataset,   
contains the maximum number of steps? `r max_steps_interval`  

## Imputing missing values
```{r}
total_missing_values <- sum(is.na(activity$steps))
```
The total number of missing values in the dataset is: `r total_missing_values`  

```{r}
new_activity <- merge(activity,stepsinterval,by="interval")
activity_filled <- mutate(new_activity, 
                          steps = case_when(
                                  is.na(steps) == TRUE ~ step_avg,
                                  TRUE ~ steps/1.0))
activity2 <- activity_filled %>% 
        select (date,interval,steps) %>% 
        arrange(date,interval)

stepsday2 <- summarize(group_by(activity2,date), 
                      step_sum = sum(steps,na.rm=TRUE))
hist(stepsday2$step_sum,xlab="Total steps per day",
     main = "Histogram of total number of steps taken each day")

steps_mean2 <- mean(stepsday2$step_sum)
steps_mean_p2 <- round(steps_mean2,1)
steps_mean_p2

steps_median2 <- median(stepsday2$step_sum)
steps_median_p2 <- round(steps_median2,1)
steps_median_p2
```  

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lubridate)
library(lattice)
activity3 <- mutate(activity2, 
                    daytype = as.factor(case_when(
                            wday(date,week_start=1) > 5 ~ "weekend",
                            TRUE ~ "weekday"))
)

stepsday3 <- summarize(group_by(activity3,daytype,interval), 
                       steps_avg = mean(steps,na.rm=TRUE))


xyplot(steps_avg  ~ interval | daytype, data= stepsday3,
       ylab = "Number of steps", xlab = "Interval",
       layout=c(1,2),type="l")


```